---
title: "Data preparation for modeling"
author : Joachim Carvallo
output: htmoutput:
  pdf_document: default
  html_document: defaultl_document
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
opts_chunk$set(warning = FALSE)
opts_chunk$set(message = FALSE)
```

```{r results = "hide}
library('dplyr') # Library for dataframes manipulations
library('ggplot2') # Library for data visualization
library('lubridate') # Library for date management
library("caret") # Library for data splitting
library("scales") # Library for better date display in plots 

# (this is how i used those libraries but it does not describe everything they can do)
```

# Creation of the database

## Weather data
```{r}
load("Donnees meteo Tours_Finales.Rda")
meteo$Desc_10h <- NULL
meteo$Desc_16h <- NULL
```

## Visitor data
```{r}
# Loading visitor data
visiteurs <- read.delim("Nombre de visiteurs par jour Villandry.txt", header = F, col.names="DATE")
visiteurs$DATE <- as.factor(visiteurs$DATE)

# Separation of the two variables
visiteurs$Nb_visiteurs <- paste(substr(visiteurs$DATE, 21, 21), substr(visiteurs$DATE, 23, 26), sep ="")
visiteurs$DATE <- substr(visiteurs$DATE, 1, 8)
visiteurs$DATE <- as.Date(visiteurs$DATE, "%d/%m/%y")
visiteurs$Nb_visiteurs <- as.numeric(visiteurs$Nb_visiteurs)

# Closing date
date_limite <- "2019-07-02"
visiteurs <- visiteurs[visiteurs$DATE<date_limite,]

# Correction of the data by tickets solded to the tourist office
place_OT <- read.delim("Places office du tourisme.txt", header = F, col.names="DATE") 
place_OT$Nb_places <- paste(substr(place_OT$DATE, 21, 21), substr(place_OT$DATE, 23, 26), sep ="")
place_OT$DATE <- substr(place_OT$DATE, 1, 8)
place_OT$DATE <- as.Date(place_OT$DATE, "%d/%m/%y")
place_OT$Nb_places <- as.numeric(place_OT$Nb_places)
place_OT <- place_OT[place_OT$DATE<date_limite,]
place_OT$Nb_places[is.na(place_OT$Nb_places)] <- 0

visiteurs$Nb_visiteurs[visiteurs$DATE >= "2010-01-01"] <- visiteurs$Nb_visiteurs[visiteurs$DATE >= "2010-01-01"] - place_OT$Nb_places
```

## Public holiday
```{r}
joursferies <- read.csv(file = "jours-feries.csv", header = T, sep = ";")
joursferies$nom_jour_ferie <- NULL
joursferies$date <- as.Date(joursferies$date, "%d/%m/%Y")
joursferies$jour_ferie <- FALSE
joursferies$jour_ferie[joursferies$est_jour_ferie == "True"] <- TRUE
joursferies$est_jour_ferie <- NULL
names(joursferies) <- c("DATE", "est_jour_ferie")
```

## Holidays
```{r}
vacances <- read.csv(file = "vacances.csv", header = T, sep = ";")
vacances$nom_vacances <- NULL
vacances$date <- as.Date(vacances$date, "%d/%m/%Y")
vacances$vacances_zone_a <- as.numeric(vacances$vacances_zone_a) -1
vacances$vacances_zone_b <- as.numeric(vacances$vacances_zone_b) -1
vacances$vacances_zone_c <- as.numeric(vacances$vacances_zone_c) -1
vacances$vacances_zone_a <- vacances$vacances_zone_a + vacances$vacances_zone_b + vacances$vacances_zone_c
vacances$vacances_zone_b <- NULL
vacances$vacances_zone_c <- NULL
names(vacances) <- c("DATE", "vacances")
```

## Fusion of the databases 
```{r}
visiteurs <- left_join(visiteurs, meteo, by = "DATE")
visiteurs <- left_join(visiteurs, joursferies, by = "DATE")
visiteurs <- left_join(visiteurs, vacances, by = "DATE")

rm(meteo, joursferies, vacances, place_OT, date_limite)
```


# Data processing 

## Missing values
```{r}
# The NA's in the number of visitors are simply days when the castle is closed (meaning 0 visitors) :
visiteurs$Nb_visiteurs[is.na(visiteurs$Nb_visiteurs)] <- 0

# Public holidays on weekends are not relevant :
visiteurs$est_jour_ferie[visiteurs$est_jour_ferie == TRUE & visiteurs$Jour %in% c(6,7)] <- FALSE

# We remove February 29 every leap year to make it easier (it's not an important day): 
visiteurs <- visiteurs[!(month(visiteurs$DATE) == 02 & day(visiteurs$DATE) == 29 & year(visiteurs$DATE) %in% c(1992,1996,2000,2004,2008,2012,2016)),]
```

## Variables Month, Day of the week and Year
```{r}
visiteurs$Mois <- as.factor(month(visiteurs$DATE))
visiteurs$Jour <- wday(visiteurs$DATE) - 1
visiteurs$Jour[visiteurs$Jour == 0] <- 7
visiteurs$Jour <- as.factor(visiteurs$Jour)
visiteurs$annee <- as.numeric(year(visiteurs$DATE))
```

## Work on the variable to be explained

The aim here is to obtain a relevant curve of the normal seasonnality in Villandry. We want to get a curve as smooth as possible, that represent, for a given day of the year, the normal frequentation (thus without knowing if it's a weekend, if there is a special event, if it's rainy, etc). Then, our variable to be explained will be the difference to the normal seasonality for each day.

```{r}
# Moving median over 7 days : 
# (moving average is too mutch dependant on the spikes of special days, it does not give a smooth curve at all)
visiteurs$mediane_gliss <- c(visiteurs$Nb_visiteurs[1:3], apply(embed(visiteurs$Nb_visiteurs, 7), 1, median), visiteurs$Nb_visiteurs[10400:10402])

# Isolation of the moving median over 27 years :
mediane_gliss <- visiteurs[visiteurs$annee %in% 1992:2018, c("DATE", "mediane_gliss")]

# Suppression of an anomaly in the data (the castle had to close a few day in 2016 because of floodings) :
mediane_gliss$mediane_gliss[mediane_gliss$DATE > "2016-05-30" & mediane_gliss$DATE < "2016-06-10"] <- c(1368.282, 1337.374, 1310.854, 1290.949, 1269.770, 1256.015, 1269.574, 1286.165, 1301.046, 1317.622)

# Averaging the moving median for each day of the year, with more weight on recent years (linear progression of weights over the years from 1 to 27) :
year(mediane_gliss$DATE) <- 2020
saison <- mediane_gliss %>% group_by(DATE) %>% summarize(Nb_visiteurs = weighted.mean(x=mediane_gliss, w=1:27+1))

# Apply a final smoothing to the curvewith a mobile average:
saison$Nb_visiteurs_liss <- c(saison$Nb_visiteurs[1:2], apply(embed(saison$Nb_visiteurs, 5), 1, mean), saison$Nb_visiteurs[364:365])

# Creation of the new variable 
visiteurs$saison <- c(rep(saison$Nb_visiteurs_liss, 28), saison$Nb_visiteurs_liss[1:182])
visiteurs$ecart_sais <- visiteurs$Nb_visiteurs - visiteurs$saison

# Plot of the resulting seasonality (before and after the last smoothing) :
ggplot(saison)+geom_line(aes(x = DATE, y = Nb_visiteurs_liss)) + theme_bw() + 
    scale_x_date(breaks = date_breaks("months"), labels = date_format("%b")) 
ggplot(saison)+geom_line(aes(x = DATE, y = Nb_visiteurs))+geom_line(aes(x = DATE, y = Nb_visiteurs_liss), col = "red") +
    theme_bw() + scale_x_date(breaks = date_breaks("months"), labels = date_format("%b")) 
# Plot of the seasonality against the year 2018
ggplot(visiteurs[visiteurs$annee == 2018,])+ geom_line(aes(x = DATE, y = Nb_visiteurs)) + 
    geom_line(aes(x = DATE, y = saison), col = "red") + theme_bw() + 
    scale_x_date(breaks = date_breaks("months"), labels = date_format("%b"))

# Animated plot of the seasonality against a few years :
library("gganimate")
temp <- visiteurs
year(temp$DATE) <- 2020
p <- ggplot(temp[temp$annee %in% c(1991,1995,2000,2005,2010,2015:2018),])+ geom_line(aes(x = DATE, y = Nb_visiteurs)) + geom_line(aes(x = DATE, y = saison), col = "red") + scale_x_date(breaks = date_breaks("months"), labels = date_format("%b")) + transition_states(annee, state_length = 3) + labs(x = "Annee: {closest_state}", y= "Nombre de visiteurs") + theme_bw(base_size = 23)

anim_save(p, file = "Animation moyenne vs annees.gif", width = 1500, height = 750, duration = 45, fps = 15)

rm(mediane_gliss, saison)

```

## New variables
```{r}
# 1) Is it a 3 day weekend ? 2) Is it the middle day of the 3 day weekend ?
visiteurs$wkd_3j <- FALSE
visiteurs$wkd_3j_milieu <- FALSE
for (i in 3:(nrow(visiteurs)-2)) {
    if (visiteurs$Jour[i] == 6 & visiteurs$est_jour_ferie[(i-1)]==TRUE) {visiteurs$wkd_3j_milieu[i] <- TRUE}
    if (visiteurs$Jour[i] == 6 & visiteurs$est_jour_ferie[(i+2)] == TRUE) {visiteurs$wkd_3j[i] <- TRUE}
    if (visiteurs$Jour[i] == 7 & visiteurs$est_jour_ferie[(i-2)]==TRUE) {visiteurs$wkd_3j[i] <- TRUE}
    if (visiteurs$Jour[i] == 7 & visiteurs$est_jour_ferie[(i+1)] == TRUE) {visiteurs$wkd_3j_milieu[i] <- TRUE}
    if (visiteurs$Jour[i] %in% c(1,5) & visiteurs$est_jour_ferie[i]==TRUE) {visiteurs$wkd_3j[i] <- TRUE}
}

# 1) Is it a 4 day weekend ? 2) Is it one of the middle days of the 4 day weekend ?
visiteurs$wkd_4j <- FALSE
visiteurs$wkd_4j_milieu <- FALSE
for (i in 4:(nrow(visiteurs)-3)) {
    if (visiteurs$Jour[i] == 6 & visiteurs$est_jour_ferie[(i-2)]==TRUE) {visiteurs$wkd_4j_milieu[i] <- TRUE}
    if (visiteurs$Jour[i] == 6 & visiteurs$est_jour_ferie[(i+3)] == TRUE) {visiteurs$wkd_4j[i] <- TRUE}
    if (visiteurs$Jour[i] == 7 & visiteurs$est_jour_ferie[(i-3)]==TRUE) {visiteurs$wkd_4j[i] <- TRUE}
    if (visiteurs$Jour[i] == 7 & visiteurs$est_jour_ferie[(i+2)] == TRUE) {visiteurs$wkd_4j_milieu[i] <- TRUE}
    if (visiteurs$Jour[i] == 5 & visiteurs$est_jour_ferie[(i-1)]==TRUE) {visiteurs$wkd_4j_milieu[i] <- TRUE}
    if (visiteurs$Jour[i] == 1 & visiteurs$est_jour_ferie[(i+1)]==TRUE) {visiteurs$wkd_4j_milieu[i] <- TRUE}
    if (visiteurs$Jour[i] %in% c(2,4) & visiteurs$est_jour_ferie[i]==TRUE) {visiteurs$wkd_4j[i] <- TRUE}
}

# Is it a "Mille feux" day ? (An event in Villandry which greatly increases attendance)
# It take place every first weekend of july since 2006 and every first weekend of august since 2018
visiteurs$mille_feux <- FALSE
visiteurs$mille_feux[visiteurs$Mois == "7" & day(visiteurs$DATE) < 8 & day(visiteurs$DATE) > 1 & visiteurs$annee>2005 & visiteurs$Jour %in% c("5", "6")] <- TRUE
visiteurs$mille_feux[visiteurs$Mois=="7" & day(visiteurs$DATE)==1 & visiteurs$annee>2005 & visiteurs$Jour==5] <- TRUE
visiteurs$mille_feux[visiteurs$Mois=="7" & day(visiteurs$DATE)==8 & visiteurs$annee>2005 & visiteurs$Jour==6] <- TRUE

visiteurs$mille_feux[visiteurs$Mois == "8" & day(visiteurs$DATE) < 8 & day(visiteurs$DATE) > 1 & visiteurs$annee>2017 & visiteurs$Jour %in% c("5", "6")] <- TRUE
visiteurs$mille_feux[visiteurs$Mois=="8" & day(visiteurs$DATE)==1 & visiteurs$DATE>"2017-12-31" & visiteurs$Jour==5] <- TRUE
visiteurs$mille_feux[visiteurs$Mois=="8" & day(visiteurs$DATE)==8 & visiteurs$DATE>"2017-12-31" & visiteurs$Jour==6] <- TRUE

# Is it one of the two major other event at Villandry ? (Heritage days and vegetable garden days)
visiteurs$wkd_patr_potag <- FALSE
visiteurs$wkd_patr_potag[visiteurs$Mois == "9" & day(visiteurs$DATE) < 22 & day(visiteurs$DATE) > 15 & visiteurs$Jour %in% c("6", "7")] <- TRUE
visiteurs$wkd_patr_potag[visiteurs$Mois == "9" & day(visiteurs$DATE) == 15 & visiteurs$Jour == 6] <- TRUE
visiteurs$wkd_patr_potag[visiteurs$Mois == "9" & day(visiteurs$DATE) == 22 & visiteurs$Jour == 7] <- TRUE

visiteurs$wkd_patr_potag[visiteurs$annee > 2005 & visiteurs$Mois == "9" & day(visiteurs$DATE) > 25 & visiteurs$Jour %in% c("6", "7")] <- TRUE
visiteurs$wkd_patr_potag[visiteurs$annee > 2005 & visiteurs$Mois == "9" & day(visiteurs$DATE) == 25 & visiteurs$Jour == 6] <- TRUE
visiteurs$wkd_patr_potag[visiteurs$annee > 2005 & visiteurs$Mois == "10" & day(visiteurs$DATE) == 1 & visiteurs$Jour == 7] <- TRUE
visiteurs$wkd_patr_potag[visiteurs$DATE %in% c("2005-09-10", "2005-09-11", "2004-09-04", "2004-09-05", "2003-09-06", "2003-09-07", "2002-09-07", "2002-09-08", "2001-09-08", "2001-09-09", "2000-09-09", "2000-09-10", "1999-09-25", "1999-09-26")] <- TRUE

# Previous days performance : 
visiteurs$ecart_prec_1 <- c(0,visiteurs$ecart_sais[1:10401])
visiteurs$ecart_prec_2 <- c(0,0,visiteurs$ecart_sais[1:10400])
visiteurs$ecart_prec_3 <- c(0,0,0,visiteurs$ecart_sais[1:10399])
visiteurs$ecart_prec_4 <- c(0,0,0,0,visiteurs$ecart_sais[1:10398])
visiteurs$ecart_prec_5 <- c(0,0,0,0,0,visiteurs$ecart_sais[1:10397])
visiteurs$ecart_prec_6 <- c(0,0,0,0,0,0,visiteurs$ecart_sais[1:10396])
visiteurs$ecart_prec_7 <- c(0,0,0,0,0,0,0,visiteurs$ecart_sais[1:10395])
```

## Last modifications :
```{r}
 # We will focus only on the periode between april and october (included) because the because for the rest of the year, the castle is often closed or receives very few visitors. The goal of the model is mostly to be used during the high season. 
visiteurs <- visiteurs[visiteurs$Mois %in% 4:10,]
visiteurs <- droplevels(visiteurs)

# Removal of the anomaly of June 2016:
visiteurs <- visiteurs[!( "2016-06-01" < visiteurs$DATE & visiteurs$DATE < "2016-06-8" ),]

# No more needed variables :
visiteurs$Nb_visiteurs <- NULL
visiteurs$mediane_gliss <- NULL
```


# Split between test set et dev set:
```{r}
# We will use 2018 and 2019 as test set in order to better visualize the predictions of our models
visiteurs_dev <- visiteurs[visiteurs$annee %in% c(2018, 2019),]
visiteurs  <- visiteurs[!(visiteurs$annee %in% c(2018, 2019)),]

visiteurs$annee <- NULL
visiteurs_dev$annee <- NULL
rm(i, p, temp)

save.image("Donnees pretes pour modelisation.Rdata")
```
