---
title: "A few statistics"
author: "Joachim CARVALLO"
date: "9 octobre 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading the database and library
```{r}
require('ggplot2')# Library for data visualization
require('lubridate') # Library for date management
require('gganimate')  # Library extending ggplot for animated plots
library('dplyr') # Library for dataframes manipulations

load("Donnees pretes pour stats.Rdata")
```


# Statistics on the number of visitors :
```{r}
# Plot of the number of visitors for 2016, 2017 and 2018
ggplot(visiteurs[visiteurs$annee == "2018",]) + 
  geom_line(aes(x=DATE, y=Nb_visiteurs), size = 1) + 
  scale_x_date(date_breaks="months", date_labels="%b") + 
  theme_bw() + labs(x = "Date", y = "Number of visitors", color = "Year")

# Change the year in the variable date in order to plot years together
visiteurs$annee <- as.factor(visiteurs$annee)
year(visiteurs$DATE) <- 2020

# Plot of the number of visitors for 2016, 2017 and 2018
ggplot(visiteurs[visiteurs$annee %in% c("2016", "2017", "2018"),]) + 
  geom_line(aes(x=DATE, y=Nb_visiteurs, color = annee), size = 1) + 
  scale_x_date(date_breaks="months", date_labels="%b") + 
  theme_bw() + labs(x = "Date", y = "Number of visitors", color = "Year")
```

```{r}
# New variable : Number of cumulative visitors
visiteurs$Nb_visiteurs_cumul_an <- 0
for (j in levels(visiteurs$annee)){
  visiteurs[visiteurs$annee == j,"Nb_visiteurs_cumul_an"][1] <- visiteurs[visiteurs$annee == j, "Nb_visiteurs"][1]
  for (i in 2:nrow(visiteurs[visiteurs$annee == j,])){
    visiteurs[visiteurs$annee == j, "Nb_visiteurs_cumul_an"][i] <- visiteurs[visiteurs$annee == j, "Nb_visiteurs"][i] + visiteurs[visiteurs$annee == j, "Nb_visiteurs_cumul_an"][(i-1)]
  }
}

# Plot of the cumulative number of visitors for 2016, 2017 and 2018
ggplot(visiteurs[visiteurs$annee %in% c("2016", "2017", "2018"),]) + 
  geom_line(aes(x=DATE, y=Nb_visiteurs_cumul_an, color = annee), size = 1) + 
  scale_x_date(date_breaks="months", date_labels="%b") + 
  theme_bw() + labs(x = "Date", y = "Number of cumulative visitors", color = "Year")
```


# Covariables against number of visitors
```{r}
# Month :
ggplot(visiteurs, aes(x=Mois, y= Nb_visiteurs)) + geom_boxplot(fill = "green3", alpha = .4) + theme_bw() + labs(x = "Month", y = "Number of visitors")

# Day of the week :
ggplot(visiteurs, aes(x=Jour, y= Nb_visiteurs)) + geom_boxplot(fill = "green3", alpha = .4) + theme_bw() + labs(x = "Day of the week", y = "Number of visitors")

# Public holidays :
ggplot(visiteurs, aes(x=est_jour_ferie, y= Nb_visiteurs)) + geom_boxplot(fill = "green3", alpha = .4) + theme_bw() + labs(x = "Public holiday ?", y = "Number of visitors")
ggplot(visiteurs, aes(x= Nb_visiteurs, fill = est_jour_ferie)) + geom_density(alpha=.5) + 
  scale_fill_manual( values = c("grey","green")) + theme_bw() + 
  labs(x = "Number of visitors", y = NULL, fill = "Public holiday ?")

# Holidays :
visiteurs$vacances <- as.factor(visiteurs$vacances)
ggplot(visiteurs, aes(x=vacances, y= Nb_visiteurs)) + geom_boxplot() + theme_bw() + labs(x = "Number of vacation zones", y = "Number of visitors")
ggplot(visiteurs, aes(x= Nb_visiteurs, fill = vacances)) + geom_density(alpha=.5) + 
  scale_fill_manual( values = c("grey","green")) + theme_bw() + 
  labs(x = "Number of visitors", y = NULL, fill = "Number of vacation zones")

# Maximum temperature :
ggplot(visiteurs, aes(x=Temp.max, y= Nb_visiteurs)) + geom_point(alpha=.1) + 
  geom_smooth(col = "green3") + theme_bw() + labs(x = "Maximum temperature of the day", y = "Number of visitors")

# Précipitations :
ggplot(visiteurs, aes(x=Precipitation+1, y= Nb_visiteurs)) + geom_point(alpha=.1) + 
  scale_x_continuous(trans = 'log2') + geom_smooth(col = "green3") + theme_bw() + 
  labs(x = "Precipitation (log scale)", y = "Deviation from seasonality of the number of visitors")

# Description of the weather at 1 pm :
ggplot(visiteurs, aes(x=Desc_13h, y= Nb_visiteurs)) + geom_boxplot(fill = "green3", alpha = .4) + theme_bw() + labs(x = "Description of the weather at 1 pm", y = "Number of visitors")

# General description of the weather :
ggplot(visiteurs, aes(x=descr_generale, y= Nb_visiteurs)) + geom_boxplot(fill = "green3", alpha = .4) + theme_bw() + labs(x = "General description of the weather", y = "Number of visitors")

# Weather grade :
ggplot(visiteurs, aes(x=note_temps, y= Nb_visiteurs)) + geom_point(alpha=.1) + 
  geom_smooth(col = "green3") + theme_bw() + labs(x = "Weather grade", y = "Number of visitors")

#  Deviation from seasonality day-1 :
ggplot(visiteurs, aes(x=ecart_prec_1, y= Nb_visiteurs)) + geom_point(alpha=.1) + 
  geom_smooth(col = "green3") + theme_bw() + labs(x = "Deviation from seasonality day-1", y = "Number of visitors")

#  Deviation from seasonality day-2 :
ggplot(visiteurs, aes(x=ecart_prec_2, y= Nb_visiteurs)) + geom_point(alpha=.1) + 
  geom_smooth(col = "green3") + theme_bw() + labs(x = "Deviation from seasonality day-2", y = "Number of visitors")

#  Deviation from seasonality day-3 :
ggplot(visiteurs, aes(x=ecart_prec_3, y= Nb_visiteurs)) + geom_point(alpha=.1) + 
  geom_smooth(col = "green3") + theme_bw() + labs(x = "Deviation from seasonality day-3", y = "Number of visitors")
```


# Covariables against deviation from seasonality

Considering the deviation from seasonality is probably more relevant here than the number of visitors because most variables are highly correlated with seasonality (temperature, weather, holidays..), which has a huge impact on the number of visitors.

```{r}
# Month :
ggplot(visiteurs, aes(x=Mois, y= ecart_sais)) + geom_boxplot(fill = "green3", alpha = .4) + theme_bw() + labs(x = "Month", y = "Deviation from seasonality of the number of visitors")

# Day of the week :
ggplot(visiteurs, aes(x=Jour, y= ecart_sais)) + geom_boxplot(fill = "green3", alpha = .4) + theme_bw() + labs(x = "Day of the week", y = "Deviation from seasonality of the number of visitors")

# Public holidays :
ggplot(visiteurs, aes(x=est_jour_ferie, y= ecart_sais)) + geom_boxplot(fill = "green3", alpha = .4) + theme_bw() + labs(x = "Public holiday ?", y = "Deviation from seasonality of the number of visitors")
ggplot(visiteurs, aes(x= ecart_sais, fill = est_jour_ferie)) + geom_density(alpha=.5) +
  scale_fill_manual( values = c("grey","green3")) + theme_bw() + 
  labs(x = "Deviation from seasonality of the number of visitors", y = NULL, fill = "Public holiday ?")

# Holidays :
visiteurs$vacances <- as.factor(visiteurs$vacances)
ggplot(visiteurs, aes(x=vacances, y= ecart_sais)) + geom_boxplot(fill = "green3", alpha = .4) + theme_bw() + labs(x = "Number of vacation zones", y = "Deviation from seasonality of the number of visitors")
ggplot(visiteurs, aes(x= ecart_sais, fill = vacances)) + geom_density(alpha=.5) + 
  scale_fill_manual( values = c("grey","green3")) + theme_bw() + 
  labs(x = "Deviation from seasonality of the number of visitors", y = NULL, fill = "Number of vacation zones")

# Maximum temperature :
ggplot(visiteurs, aes(x=Temp.max, y= ecart_sais)) + geom_point(alpha=.1) + 
  geom_smooth(col = "green3") + theme_bw() + labs(x = "Maximum temperature of the day", y = "Deviation from seasonality of the number of visitors")

# Précipitations :
ggplot(visiteurs, aes(x=Precipitation+1, y= ecart_sais)) + geom_point(alpha=.1) +
  scale_x_continuous(trans = 'log10') + geom_smooth(col = "green3") + theme_bw() + 
  labs(x = "Precipitation (log scale)", y = "Deviation from seasonality of the number of visitors")

# Description of the weather at 1 pm :
ggplot(visiteurs, aes(x=Desc_13h, y= ecart_sais)) + geom_boxplot(fill = "green3", alpha = .4) + theme_bw() + labs(x = "Description of the weather at 1 pm", y = "Deviation from seasonality of the number of visitors")

# General description of the weather :
ggplot(visiteurs, aes(x=descr_generale, y= ecart_sais)) + geom_boxplot(fill = "green3", alpha = .4) + theme_bw() + labs(x = "General description of the weather", y = "Deviation from seasonality of the number of visitors")

# Weather grade :
ggplot(visiteurs, aes(x=note_temps, y= ecart_sais)) + geom_point(alpha=.1) + 
  geom_smooth(col = "green3") + theme_bw() + labs(x = "Weather grade", y = "Deviation from seasonality of the number of visitors")

#  Deviation from seasonality day-1 :
ggplot(visiteurs, aes(x=ecart_prec_1, y= ecart_sais)) + geom_point(alpha=.1) + 
  geom_smooth(col = "green3") + theme_bw() + labs(x = "Deviation from seasonality day-1", y = "Number of visitors")
```


# Animations
```{r}
# Number of cumulative visitors for 2016, 2017 and 2018
p <- ggplot(visiteurs[visiteurs$annee %in% c("2016", "2017", "2018"),], aes(x=DATE, y=Nb_visiteurs_cumul_an, color = annee)) +
  geom_line(size = 0.9) + geom_text(aes(label = gsub(pattern = "20", replacement = "", x = annee)), size=12, hjust = 0) + 
  scale_x_date(date_breaks="months", date_labels="%b") + 
  theme_bw(base_size = 30) + labs(color="Year", y= "Number of cumulative visitors", x = "Date") +
  theme(legend.position="top", plot.margin = margin(r=120, l=20, t=20, b=75)) + 
  transition_reveal(DATE) + view_follow()

anim_save(p, file = "Number of cumulative visitors for 2016 2017 2018.gif", width = 750, height = 500, res = 40, duration = 25, start_pause = 24, end_pause = 120, fps = 24)
```

```{r}
# Test animation :
temp <- visiteurs[visiteurs$annee %in% as.character(seq(2005,2019,1)),]
temp <- temp[!is.na(temp$DATE),]

liste <- data.frame(Nb_visiteurs_cumul_an=c(), annee=c(), DATE=c(), classement=c())
for (jour in temp$DATE[temp$annee == "2019"]){
  nb_vis <- temp[temp$DATE == jour, c("Nb_visiteurs_cumul_an", "annee"),]
  nb_vis <- nb_vis[!is.na(nb_vis$Nb_visiteurs_cumul_an),]
  nb_vis <- nb_vis[order(nb_vis$Nb_visiteurs_cumul_an),]
  nb_vis$DATE <- as.Date(jour, origin = as.Date("1970-01-01"))
  nb_vis$classement <- seq(1,nrow(nb_vis),1)
  liste <- rbind(liste, nb_vis)
}
temp <- left_join(temp, liste, by = c("DATE", "Nb_visiteurs_cumul_an", "annee"))
rm(liste, jour, nb_vis)

# Theme
my_background <- 'antiquewhite'
my_pal <- "viridis" #colors for bars (from wesanderson)
my_theme <- my_theme <- theme(rect = element_rect(fill = my_background),
                              plot.background = element_rect(fill = my_background, color = NA),
                              panel.background = element_rect(fill = my_background, color = NA),
                              panel.border = element_blank(),
                              plot.title = element_text(face = 'bold', size = 30),
                              panel.grid.major.y = element_blank(),
                              panel.grid.minor.y = element_blank(),
                              panel.grid.major.x = element_line(color = 'grey75'),
                              panel.grid.minor.x = element_line(color = 'grey75'),
                              legend.position = 'none',
                              plot.caption = element_text(size = 20),
                              axis.ticks = element_blank(),
                              axis.text.y =  element_blank())

theme_set(theme_light() + my_theme)

p <- ggplot(aes(classement, group = annee), data = temp) +
  geom_tile(aes(y = Nb_visiteurs_cumul_an / 2, 
                height = Nb_visiteurs_cumul_an,
                width = 0.9, fill = as.numeric(as.character(annee))), alpha = 0.9) + scale_fill_viridis_c() +
  geom_text(aes(y = Nb_visiteurs_cumul_an+10000, label = annee), nudge_y = -2, size = 15) +
  #convert to character to get rid of blurry trailing decimals
  geom_text(aes(x=17,y=175000, label=paste0(DATE)), size=20, color = 'gray45') +
  scale_y_continuous(breaks=c(0, 50000, 100000, 150000, 200000, 250000, 300000)) +
  coord_cartesian(clip = "off", expand = FALSE) +
  coord_flip() +
  labs(title = 'Best years :',
       x = '',
       y = '') +
  transition_states(DATE, transition_length = 3, state_length = 2) +
  ease_aes('cubic-in-out')


anim_save(p, file = "Animation visiteurs_Classement années.gif", width = 1500, height = 1000, res = 40, duration = 120, start_pause = 48, end_pause = 240, fps = 24)
```
