---
title: "Optimization and analysis of the model"
author: "Joachim CARVALLO"
date: "9 décembre 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading libraries and data
```{r}
library('caret') # Library for dummy variables and RMSE
library('xgboost') # Library for Xgboost
library('lubridate') # Library for date management
library('ggplot2') # Library for data visualization
library('dplyr') # Library for dataframe manipulations
library("SHAPforxgboost") # Library implementing Shap value

# Loading data
load("Donnees pretes pour modelisation.Rdata")

# Factor variables to dummy variables
DATE <- visiteurs_test$DATE
dmy <- dummyVars( ~. , data = visiteurs[,-1], fullRank = T)
visiteurs <- data.frame(predict(dmy, newdata = visiteurs))
dmy <- dummyVars( ~. , data = visiteurs_test[,-1], fullRank = T)
visiteurs_test <- data.frame(predict(dmy, newdata = visiteurs_test))
visiteurs$descr_generalePas.d.obs <- NULL # This doesn't occur in the test set 

# Data in the right format
dtrain <- xgb.DMatrix(data = as.matrix(visiteurs[,-33]), label=as.matrix(visiteurs$ecart_sais))
dtest <- xgb.DMatrix(data = as.matrix(visiteurs_test[,-33]), label=as.matrix(visiteurs_test$ecart_sais))
watchlist <- list(train=dtrain, test=dtest)
```


# Optimization of the model 

## First round 
```{r}
# Random grid of parameters :
n <- 300 # size of the grid
random_grid <- data.frame(
                          eta = 0.02,
                          gamma = 10^(runif(n, min = -3, max = 2)),
                          max_depth = sample(1:12, size = n, replace = T),
                          subsample = runif(n, min = 0.3, max = 1),
                          colsample_bytree = runif(n, min = 0.3, max = 1),
                          min_child_weight = 10^(runif(n, min = -3, max = 2)),
                          RMSE = 0
)

for(i in 1:n){
  params <- as.list(random_grid[i,])
  
  # Training :
  xgb <- xgb.train(params = params, data = dtrain, watchlist=watchlist, objective = "reg:squarederror", 
                   verbose = 0, nthread = 3, nrounds = 10000, early_stopping_rounds = 250)
  random_grid$RMSE[i] <- xgb$best_score
  print(i)
}

#save(random_grid, file = "Grille d'optimisation des paramètres du XGBoost_1.Rdata")
```

Best model :

| Parameter 	     |  Value      |
| ---------------- |:-----------:|
| eta              | 0.02        |
| gamma            | 0.513787589 |
| max_depth        | 8           |
| subsample        | 0.4299943   |
| colsample_bytree | 0.4935486   |
| min_child_weight | 4.141807383 |
| *RMSE*           | *241.5967*  |

## Second round 
```{r}
# New random grid of parameters :
n <- 300 # size of the grid
random_grid_2 <- data.frame(
                          eta = 0.01,
                          gamma = 10^(runif(n, min = -1.5, max = 0.5)),
                          max_depth = sample(6:10, size = n, replace = T),
                          subsample = runif(n, min = 0.35, max = .7),
                          colsample_bytree = runif(n, min = 0.4, max = .6),
                          min_child_weight = 10^(runif(n, min = -0.5, max = 1.5)),
                          RMSE = 0
)

for(i in 1:n){
  params <- as.list(random_grid_2[i,])
  # Training :
  xgb <- xgb.train(params = params, data = dtrain, watchlist=watchlist, objective = "reg:squarederror", 
                   verbose = 0, nthread = 3, nrounds = 10000, early_stopping_rounds = 250)
  random_grid_2$RMSE[i] <- xgb$best_score
  print(i)
}

#save(random_grid_2, file = "Grille d'optimisation des paramètres du XGBoost_2.Rdata")
```

Best model :

| Parameter 	     |  Value      |
| ---------------- |:-----------:|
| eta              | 0.01        |
| gamma            | 0.74593472  |
| max_depth        | 7           |
| subsample        | 0.4751654   |
| colsample_bytree | 0.5389854   |
| min_child_weight | 6.2127380   |
| *RMSE*           | *240.0006*  |

## Final model 
```{r}
params <- as.list(random_grid_2[which.min(random_grid_2$RMSE), 1:6])

# Final training :
xgb <- xgb.train(params = params, data = dtrain, watchlist=watchlist, objective = "reg:squarederror", 
                   verbose = 2, nthread = 3, nrounds = 10000, early_stopping_rounds = 250)

# RMSE vs number of iterations :
ggplot(xgb$evaluation_log) + geom_line(aes(x = iter, y= test_rmse)) + 
  geom_line(aes(x = iter, y= train_rmse), color = "red") + theme_bw()

# Save final model 
#save(xgb, file = "C:/Users/joach/OneDrive/Bureau/Visites Villandry/XGBoost prevision freq Villandry - V3.Rdata")
```


# Error analysis 
```{r}
load("Donnees pretes pour modelisation.Rdata")

# Make prediction on test set and compute RMSE
visiteurs_test$predict <- predict(xgb, dtest)

# Plot of the real attendance against predicted for 2016
ggplot(visiteurs_test[year(visiteurs_test$DATE)==2016,]) + 
  geom_line(aes(x = DATE, y = predict + saison), col="green3") + 
  geom_line(aes(x = DATE, y = ecart_sais + saison)) +  
  theme_bw() +labs(x = "Date (2018)", y = "Number of visitors : prediction vs real")

# Répartition des écarts : 
ggplot(visiteurs_test) + geom_density(aes(x = ecart_sais - predict), fill = "grey") + theme_bw()

# Percentage distribution of errors : (density)
visiteurs_test$prop <- (visiteurs_test$predict - visiteurs_test$ecart_sais)*100/(visiteurs_test$saison + visiteurs_test$ecart_sais)
ggplot(visiteurs_test) + geom_density(aes(x = prop), fill = "grey") + theme_bw()

mean(abs(visiteurs_test$prop) < 7) # 33,3% du temps 
mean(abs(visiteurs_test$prop) < 17) # 66,9% du temps 
mean(abs(visiteurs_test$prop) < 27) # 84,6% du temps 
mean(abs(visiteurs_test$prop) < 40) # 95% du temps

# Percentage distribution of errors : (bar plot)
prop <- c(mean(abs(visiteurs_test$prop)<10), mean(abs(visiteurs_test$prop)>10 & abs(visiteurs_test$prop)<20),
  mean(abs(visiteurs_test$prop)>20 & abs(visiteurs_test$prop)<30), 
  mean(abs(visiteurs_test$prop)>30 & abs(visiteurs_test$prop)<40),
  mean(abs(visiteurs_test$prop)>40 & abs(visiteurs_test$prop)<60), 
  mean(abs(visiteurs_test$prop)>60 & abs(visiteurs_test$prop)<80),mean(abs(visiteurs_test$prop)>80))

temp <- data.frame(prop =  prop, prop_erreur = c("< 10%","< 20%","< 30%","< 40%","< 60%", "< 80%", "> 80%"))

ggplot(temp) + 
  geom_bar(aes(x = prop_erreur, y = prop), stat = "identity", fill = "green3", alpha = .5, col = "black") +
  scale_y_continuous(breaks = c("40%"=0.4, "30%"=0.3, "20%"=0.2, "10%"=0.1, "3%"=0.03)) + theme_bw() + 
  labs(x = "Error as a proportion of the number of visitors", y = "Volume")
```




# Analysis of the model

## SHAP importance and dependence plot
```{r}
load("C:/Users/joach/OneDrive/Bureau/Visites Villandry/XGBoost prevision freq Villandry - V3.Rdata")


### SHAP importance plot :

temp <- shap_long %>% group_by(variable) %>% summarise(mean_value = mean_value[1])
ggplot(temp[temp$mean_value > 7,]) + 
  geom_bar(aes(x = reorder(variable, mean_value), y = mean_value), stat = "identity", fill = "green3", alpha = .5) +
  coord_flip() + theme_bw() + labs(x = "Variables", y = "Mesure of the importance of variables")

### SHAP dependence plot :

# Difference from the seasonality of the day before :
ggplot(shap_long[shap_long$variable == "ecart_prec_1",], aes(x = rfvalue, y = value)) + 
  geom_point(cex = 1, alpha = .3) + geom_smooth(method = "loess", col = "green3") + 
  theme_bw() + labs(x = "Difference from the seasonality of the day before", y = "SHAP value")
# Sundays :
ggplot(shap_long[shap_long$variable == "Jour.7",], aes(x = as.logical(rfvalue), y = value)) + 
  geom_boxplot(fill = "green3", alpha = .5) + theme_bw() + labs(x = "Sunday ?", y = "SHAP value")
# Weather grade :
ggplot(shap_long[shap_long$variable == "note_temps",], aes(x = rfvalue, y = value)) + 
  geom_point(cex = 1, alpha = .3) + geom_smooth(method = "loess", col = "green3") + 
  theme_bw() + labs(x = "Weather grade", y = "SHAP value")
# Difference from the seasonality : day - 7 :
ggplot(shap_long[shap_long$variable == "ecart_prec_7",], aes(x = rfvalue, y = value)) + 
  geom_point(cex = 1, alpha = .3) + geom_smooth(method = "loess", col = "green3") + 
  theme_bw() + labs(x = "Difference from the seasonality : day - 7", y = "SHAP value")
# Precipitation :
ggplot(shap_long[shap_long$variable == "Precipitation",], aes(x = rfvalue+1, y = value)) + 
  geom_point(cex = 1, alpha = .3) + geom_smooth(method = "loess", col = "green3") + scale_x_continuous(trans = "log10") +
  theme_bw() + labs(x = "Precipitation", y = "SHAP value")
# Maximum temperature :
ggplot(shap_long[shap_long$variable == "Temp.max",], aes(x = rfvalue, y = value)) + 
  geom_point(cex = 1, alpha = .3) + geom_smooth(method = "loess", col = "green3") + 
  theme_bw() + labs(x = "Maximum temperature", y = "SHAP value")
# Middle of a 4 days weekend :
ggplot(shap_long[shap_long$variable == "wkd_4j_milieuTRUE",], aes(x = as.logical(rfvalue), y = value)) + 
  geom_boxplot(fill = "green3", alpha = .5) + theme_bw() + labs(x = "Middle of a 4-day weekend ?", y = "SHAP value")
```

## SHAP interaction effect plot
```{r}
# Maximum temperature :
variable <- "Temp.max"

intercations_moy <- apply(abs(shap_int[,,variable]), 2, mean)
interaction_1 <- names(which.max(intercations_moy[names(intercations_moy) != variable]))
interaction_2 <- names(which.max(intercations_moy[!(names(intercations_moy) %in% c(variable, interaction_1))]))

temp <- data.frame(rfvalue = visiteurs[,variable], 
                   main_effect = shap_int[,variable,variable], 
                   interaction_1 = shap_int[,interaction_1,variable], 
                   interaction_2 = shap_int[,interaction_2,variable],
                   rfvalue_interaction_1 = visiteurs[,interaction_1], 
                   rfvalue_interaction_2 = visiteurs[,interaction_2])

ggplot(temp, aes(x = rfvalue, y = main_effect)) + geom_point(cex = 1, alpha = .3) + 
  geom_smooth(method = "loess", col = "green3") +  theme_bw() + 
  labs(x = "Maximum temperature", y = "SHAP value - Main effect")

ggplot(temp, aes(x = rfvalue, y = interaction_1, col = rfvalue_interaction_1+1)) + 
  geom_point(cex = 1, alpha = .3) + scale_color_viridis_c(trans = "log10") + theme_bw() + 
  labs(x = "Maximum temperature", y = "SHAP value - Interaction with precipitations", col = "Precipitations")
ggplot(temp, aes(x = rfvalue, y = main_effect+interaction_1, col = rfvalue_interaction_1+1)) + 
  geom_point(cex = 1, alpha = .3) + scale_color_viridis_c(trans = "log10") + theme_bw() + 
  labs(x = "Maximum temperature", y = "SHAP value - Interaction with precipitations", col = "Precipitations")

ggplot(temp, aes(x = rfvalue, y = interaction_2, col = rfvalue_interaction_2)) + 
  geom_point(cex = 1, alpha = .3) + scale_color_viridis_c() + theme_bw() + 
  labs(x = "Maximum temperature", y = "SHAP value - Interaction with weather grade", col = "Weather grade")
ggplot(temp, aes(x = rfvalue, y = main_effect+interaction_2, col = rfvalue_interaction_2)) + 
  geom_point(cex = 1, alpha = .3) + scale_color_viridis_c() + theme_bw() + 
  labs(x = "Maximum temperature", y = "SHAP value - Interaction with weather grade", col = "Weather grade")


# Precipitations :
variable <- "Precipitation"

intercations_moy <- apply(abs(shap_int[,,variable]), 2, mean)
interaction_1 <- names(which.max(intercations_moy[names(intercations_moy) != variable]))
interaction_2 <- names(which.max(intercations_moy[!(names(intercations_moy) %in% c(variable, interaction_1))]))

temp <- data.frame(rfvalue = visiteurs[,variable], main_effect = shap_int[,variable,variable], 
                   interaction_1 = shap_int[,interaction_1,variable], 
                   interaction_2 = shap_int[,interaction_2,variable], 
                   rfvalue_interaction_1 = visiteurs[,interaction_1], 
                   rfvalue_interaction_2 = visiteurs[,interaction_2])

ggplot(temp, aes(x = rfvalue, y = main_effect)) + geom_point(cex = 1, alpha = .3) + 
  geom_smooth(method = "loess", col = "green3") + theme_bw() + 
  labs(x = "Maximum temperature", y = "SHAP value - Main effect")

ggplot(temp, aes(x = rfvalue+1, y = interaction_1, col = rfvalue_interaction_1+1)) + 
  geom_point(cex = 1, alpha = .3) + scale_color_viridis_c() + scale_x_continuous(trans = "log10") + 
  theme_bw() + labs(x = "Precipitations", y = "SHAP value - Interaction with weather grade", col = "Weather grade")
ggplot(temp, aes(x = rfvalue+1, y = main_effect+interaction_1, col = rfvalue_interaction_1+1)) + 
  geom_point(cex = 1, alpha = .3) + scale_color_viridis_c() + scale_x_continuous(trans = "log10") + theme_bw() + 
  labs(x = "Precipitations", y = "SHAP value - Interaction with weather grade", col = "Weather grade")

ggplot(temp, aes(x = rfvalue+1, y = interaction_2, col = rfvalue_interaction_2)) + 
  geom_point(cex = 1, alpha = .3) + scale_x_continuous(trans = "log10") + scale_color_viridis_c() + theme_bw() + 
  labs(x = "Precipitations", y = "SHAP value - Interaction with seasonality", col = "Seasonality")
ggplot(temp, aes(x = rfvalue+1, y = main_effect+interaction_2, col = rfvalue_interaction_2)) + 
  geom_point(cex = 1, alpha = .3) + scale_x_continuous(trans = "log10") + scale_color_viridis_c() + theme_bw() + 
  labs(x = "Precipitations", y = "SHAP value - Interaction with seasonality", col = "Seasonality")
```

## Water fall
```{r}
threshold = 10
data = visiteurs_test[81,]

dtest <- xgb.DMatrix(data = as.matrix(data[-33]), label=as.matrix(data$ecart_sais))
explain_obs <- predict(xgb, dtest, predcontrib = TRUE, approxcontrib = FALSE)

# Get data
breakdown_summary = explain_obs[1,-46]
weight = rowSums(explain_obs)
labels = names(breakdown_summary)
data_for_label = round(as.matrix(data[-33]), 1)
Biais = explain_obs[46]

# Order on contribution value
i = order(abs(breakdown_summary), decreasing = TRUE)
breakdown_summary = breakdown_summary[i]
data_for_label = data_for_label[i]

# Regroup variables with impact < threshold
i_other = which(abs(breakdown_summary) < threshold)
other_impact = 0
other_impact = sum(breakdown_summary[i_other])
names(other_impact) = "other"
breakdown_summary = breakdown_summary[-i_other]
data_for_label = data_for_label[-i_other]

# Regroup everything
breakdown_summary = c(Biais, breakdown_summary, other_impact)
data_for_label = c("", data_for_label, "")
labels = paste0(names(breakdown_summary), " = ", data_for_label)
labels[1] = "Biais"
labels[length(labels)] = "Other"

# Defini colors 
fill_colours  <- rep("green3", length(breakdown_summary))
fill_colours[sign(breakdown_summary) == -1] <- "grey"

# Plot the waterfall
waterfalls::waterfall(values = breakdown_summary, 
                      rect_text_labels = round(breakdown_summary, 0), 
                      labels = labels, 
                      total_rect_text = round(weight, 0), 
                      calc_total = TRUE, 
                      total_axis_text = "Prediction", 
                      fill_by_sign = F, 
                      fill_colours = fill_colours) + 
  
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        panel.background = element_rect("white", "black"),
        panel.grid.major = element_line("grey"), 
        panel.grid.minor = element_line(color = "grey", size = .2)) +
  
  labs(title = paste(c("Résultat = Saisonnalité + Prédiction =", 
                       as.character(round(data$saison)), 
                       "+", 
                       as.character(round(sum(explain_obs))), 
                       "=", 
                       as.character(round(data$saison+sum(explain_obs)))), 
        collapse = " "))
```
